Bird beak's elaboration and innovation

### 2022-03-01 Analysis/10-Figure1.Rmd
Update 10-Figure1.Rmd

    ## Do the empty plot
    plot(NULL, xlim = c(-1.25, 1.25), ylim = c(-1.25, 1.25), xaxt = "n", yaxt = "n", xlab = ifelse(x, paste0("PC1 (", pc_var[1], "%)"), ""), ylab = ifelse(y, paste0("PC2 (", pc_var[2], "%)"), ""), main = main)
    if(x) axis(1)
    if(y) axis(2)
    abline(v = 0, col = "grey", lwd = 0.5)
    abline(h = 0, col = "grey", lwd = 0.5)

    if(!with.phylo) {
        subsets <- names(name.col)
        cols <- name.col
    } else {
        subsets <- c("phylogeny", names(name.col))
        cols <- c("grey", name.col)
    }
    covar.plot(get.subsets(data, subsets = subsets),
           col = cols,
           ellipses = mean,
           legend = FALSE,
           points = FALSE,
           cex = 0.5, apply.to.VCV = TRUE,
           lwd = 4,
           add = TRUE,
           legend = legend)
}
## Plot the projection/rejection
plot.one.proj.rej <- function(data, name.col, x = TRUE, y = TRUE, main = "") {

@@

    ## Plot the results
    boxplot(cbind(elaboration, exploration), horizontal = TRUE, ylim = c(0,lim_up), col = col, las = 1, xaxt = "n", yaxt = "n")
    if(x) {
        axis(1, at = seq(from = 0, to = lim_up, by = 5), labels = seq(from = 0, to = lim_up, by = 5))
    }
    if(y) {
        axis(2, at = c(1, 2), labels = c("exploration", "innovation"), las = 2)
    }
}

@@

@@

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Paleognathae" = 1,
    "Galloanserae" = 2,
    "Otidimorphae" = 5,
    "Aequornithes" = 6,
    "Mirandornithes" = 3,
    "Columbimorphae" = 4,
    "Telluraves" = 8,
    "Strisores" = 7)
tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))
# par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
# plot(rep(1,27), 1:27, col = "white")
# labs <- names(tip_colours)
# cols <- tip_colours
# text(x = rep(1,27), y = 1:27, label = rev(labs), col = rev(cols))

for(i in 1:length(tip_order)) {
    ## Select the colour and name
    name.col <- tip_colours[match(sort(names(tip_colours)), names(tip_colours))][i]

    ## Plot the ellipse
    par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
    plot.one.ellipse(superorder_shapespace, name.col, x = FALSE, y = FALSE, with.phylo = TRUE)
    text(0,1.2, names(name.col), cex = 1.5, col = name.col)

@@

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Tinamiformes"= 1,
    "Galliformes"= 3,
    "Anseriformes"= 2,
    "Otidiformes"= 7,
    "Cuculiformes" = 8,
    "Gruiformes" = 9, 
    "Musophagiformes" = 10,
    "Procellariiformes" = 12,
    "Sphenisciformes" = 11,
    "Ciconiiformes" = 13,
    "Pelecaniformes"= 15,
    "Suliformes" = 14,
    "Podicipediformes" = 4,
    "Columbiformes" = 6,
    "Pterocliformes" = 5,
    "Charadriiformes" = 18,
    "Trogoniformes" = 21,
    "Bucerotiformes" = 22,
    "Coraciiformes" = 23,
    "Piciformes" = 24,
    "Strigiformes" = 20,
    "Accipitriformes" = 19, 
    "Falconiformes" = 25,
    "Psittaciformes" = 26,
    "Passeriformes" = 27,
    "Caprimulgiformes" = 16, 
    "Apodiformes" = 17)
tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))

for(i in 1:length(tip_order)) {
    ## Select the colour and name
    name.col <- tip_colours[match(sort(names(tip_colours)), names(tip_colours))][i]

    ## Plot the ellipse
    par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
    plot.one.ellipse(order_shapespace, name.col, x = FALSE, y = FALSE, with.phylo = TRUE)
    text(0,1.2, names(name.col), cex = 1.5, col = name.col)

>>>>

    ## Do the empty plot
    lims <- c(-1/3, 1/3) # c(-1.25, 1.25)
    plot(NULL, xlim = lims, ylim = lims, xaxt = "n", yaxt = "n", xlab = ifelse(x, paste0("PC1 (", pc_var[1], "%)"), ""), ylab = ifelse(y, paste0("PC2 (", pc_var[2], "%)"), ""), main = main)
    if(x) axis(1)
    if(y) axis(2)
    abline(v = 0, col = "grey", lwd = 0.5)
    abline(h = 0, col = "grey", lwd = 0.5)

    if(!with.phylo) {
        subsets <- names(name.col)
        cols <- name.col
    } else {
        subsets <- c("phylogeny", names(name.col))
        cols <- c("grey", name.col)
    }
    if(!with.phylo) {
        covar.plot(get.subsets(data, subsets = subsets),
               col = cols,
               ellipses = mean,
               legend = FALSE,
               points = FALSE,
               cex = 0.5, apply.to.VCV = TRUE,
               lwd = 4,
               add = TRUE,
               legend = legend)
    } else {
        covar.plot(get.subsets(data, subsets = subsets),
               col = cols,
               ellipses = mean,
               legend = FALSE,
               points = FALSE,
               cex = 0.5, apply.to.VCV = TRUE,
               lwd = 4,
               add = TRUE,
               legend = legend,
               scale = "phylogeny")        
    }
}
## Plot the projection/rejection
plot.one.proj.rej <- function(data, name.col, x = TRUE, y = TRUE, main = "") {

@@

    ## Get the quantiles
    elaboration_quantiles <- quantile(elaboration, prob = c(0.025, 0.25, 0.50, 0.975))
    exploration_quantiles <- quantile(exploration, prob = c(0.025, 0.25, 0.50, 0.975))

    ## Empty plot
    par(bty = "n")
    plot(NULL, xlim = c(0,1.5), ylim = c(0,lim_up), xaxt = "n", yaxt = "n", xlab = "", ylab = "")

    ## Add the lines
    lines(lty = 2, x = c(0,0), y = elaboration_quantiles[c(1,4)], col = col[1], lwd = 1)
    lines(lty = 1, x = c(0,0), y = elaboration_quantiles[c(2,3)], col = col[1], lwd = 3)
    lines(lty = 2, x = c(1,1), y = exploration_quantiles[c(1,4)], col = col[2], lwd = 1)
    lines(lty = 1, x = c(1,1), y = exploration_quantiles[c(2,3)], col = col[2], lwd = 3)
    points(0, median(elaboration), pch = 19, col = col[1])
    points(1, median(exploration), pch = 19, col = col[2])

    ## Add the y axis
    axis(4, at = seq(from = 0, to = lim_up, by = 5), labels = seq(from = 0, to = lim_up, by = 5), las = 2)

    ## Plot the results
    # boxplot(cbind(elaboration, exploration), horizontal = TRUE, ylim = c(0,lim_up), col = col, las = 1, xaxt = "n", yaxt = "n")
    # if(x) {
    #     axis(1, at = seq(from = 0, to = lim_up, by = 5), labels = seq(from = 0, to = lim_up, by = 5))
    # }
    # if(y) {
    #     axis(2, at = c(1, 2), labels = c("exploration", "innovation"), las = 2)
    # }
}

@@

# sequence <- c(c(rep(2, 3), 3), c(rep(4, 3), 5), c(rep(6, 3), 7), c(rep(8, 3), 9))

@@

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Paleognathae" = 1,
    "Galloanserae" = 2,
    "Otidimorphae" = 5,
    "Aequornithes" = 6,
    "Mirandornithes" = 3,
    "Columbimorphae" = 4,
    "Telluraves" = 8,
    "Strisores" = 7)

tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))

# par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
# plot(rep(1,27), 1:27, col = "white")
# labs <- names(tip_colours)
# cols <- tip_colours
# text(x = rep(1,27), y = 1:27, label = rev(labs), col = rev(cols))

for(i in 1:length(tip_order)) {
    ## Select the colour and name
    # name.col <- tip_colours[match(sort(names(tip_colours)), names(tip_colours))][i]
    name.col <- tip_colours[i]

    ## Plot the ellipse
    par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
    plot.one.ellipse(superorder_shapespace, name.col, x = FALSE, y = FALSE, with.phylo = TRUE)
    text(0,0.3, names(name.col), cex = 1.5, col = name.col)

@@

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Tinamiformes"= 1,
    "Galliformes"= 3,
    "Anseriformes"= 2,
    "Otidiformes"= 7,
    "Cuculiformes" = 8,
    "Gruiformes" = 9, 
    "Musophagiformes" = 10,
    "Procellariiformes" = 12,
    "Sphenisciformes" = 11,
    "Ciconiiformes" = 13,
    "Pelecaniformes"= 15,
    "Suliformes" = 14,
    "Podicipediformes" = 4,
    "Columbiformes" = 6,
    "Pterocliformes" = 5,
    "Charadriiformes" = 18,
    "Trogoniformes" = 21,
    "Bucerotiformes" = 22,
    "Coraciiformes" = 23,
    "Piciformes" = 24,
    "Strigiformes" = 20,
    "Accipitriformes" = 19, 
    "Falconiformes" = 25,
    "Psittaciformes" = 26,
    "Passeriformes" = 27,
    "Caprimulgiformes" = 16, 
    "Apodiformes" = 17)

# tip_order <- sort(tip_order)

tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))

for(i in 1:length(tip_order)) {
    ## Select the colour and name
    # name.col <- tip_colours[match(sort(names(tip_colours)), names(tip_colours))][i]
    name.col <- tip_colours[i]

    ## Plot the ellipse
    par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
    plot.one.ellipse(order_shapespace, name.col, x = FALSE, y = FALSE, with.phylo = TRUE)
    text(0,0.3, names(name.col), cex = 1.5, col = name.col)
